Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

### 2025-01-21 manuscript/appendix.qmd
Mega revisions (almost done!)

---
title: 'Appendix for "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"'
short-title: 'Appendix for "Why Donors Donate"'
date: "2025-01-11"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
toc: true
toc-depth: 2
format: 
  hikmah-pdf:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
  hikmah-manuscript-pdf+ms:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
---

@@

   :   We rephrased this as H~1b~: "If NGOs are accountable and hold regular third-party audits, then individual private donors will have a higher likelihood of supporting or donating to them."

>>>>

---
title: 'Appendix for "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"'
short-title: 'Appendix for "Why Donors Donate"'
date: "2025-01-20"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
toc: true
toc-depth: 2
format: 
  hikmah-pdf:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
  hikmah-manuscript-pdf+ms:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
---

@@

   :   We rephrased this as H~1b~: "If NGOs are accountable, then individual private donors will have a higher likelihood of supporting or donating to them."

### 2025-01-21 manuscript/manuscript.qmd
Mega revisions (almost done!)

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2025-01-17"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
correspondence-prefix: "Correspondence concerning this article should be addressed to"
abstract: |
  Governments across the world have increasingly used laws to restrict the work of nonprofits, which has led to a reduction in public or official foreign aid directed towards these groups. Many international nonprofits, in response, have turned to individual donors to offset the loss of traditional fundings. What are individual donors' preferences regarding donating to legally besieged nonprofits abroad? We conducted a conjoint experiment on a nationally representative sample of likely donors in the US and found that learning about host government criticism and restrictions on nonprofits decreases individuals' preferences to donate to them. However, organizational features such as financial transparency and accountability can protect against this dampening effect. Our results have important implications both for understanding private international philanthropy and how nonprofits can better frame their fundraising appeals at a time when they are facing restrictive civic spaces and hostile governments abroad.
additional-info: |
  **Data Availability:** All data and replication code are publicly available at <https://github.com/andrewheiss/silent-skywalk>
keywords:
  - philanthropy
  - conjoint experiments
  - donor heuristics
  - repression
  - NGOs
  - civil society
  - nonprofits
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
word-count: "9,300"
---

@@

In the remainder of the paper, we provide an overview of current research on private philanthropic giving to international NGOs. We use these existing theories to hypothesize how organizational practices like financial transparency and accountability influence individual donor behavior in an environment of closing civic space. We then detail our conjoint study methodology and discuss our findings. Finally, we use our results to offer insights for NGOs facing legal restriction abroad.

@@

Though individual donors would like to have some assurance that their resources will be used appropriately and organizations are not spending too much on overhead, individuals interested in supporting charities cannot thoroughly vet every aspect of the organization [@Tremblay-BoirePrakash:2017; @ChaudhryHeiss:2021]. To more easily judge financial transparency and accountability, donors may use ratings by watchdog organizations like Charity Navigator and GuideStar as heuristics to guide their decisions.[^cn-multiple] Research shows that initiatives to address information gaps by these organizations can substantially increase donations [@Sloan:2009; @GordonKnockNeely:2009; @Tremblay-BoirePrakash:2017].

@@

However, previous literature has established that individual private donors do not make the same considerations as official donors when deciding to donate [@DesaiKharas:2018; @ChaudhryHeiss:2021]. Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. When a potential recipient NGO faces legal crackdown in its host country, private individual donors can respond by (1) increasing their donations as a sign of support and solidarity, (2) decreasing their donations, punishing the NGO for doing something to run afoul of its host government, or (3) not considering the host country legal environment and making no change in their donation behavior. @ChaudhryDotsonHeiss:2021 explore how different combinations of donor characteristics determine which response individuals are likely to take when NGOs face crackdown. They find that donors with longer experiences with the nonprofit sector and high levels of social trust---i.e. those who frequently volunteer, regularly donate to charity, and trust political institutions---are more likely to maintain their support for international NGOs that face criticism or crackdown abroad. Thus, anti-NGO laws may act as a heuristic to individuals donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and seek to restrict their work.

@@

> **H~2~**: If NGOs face legal crackdowns abroad, then individual private donors will have a higher likelihood of supporting or donating to them.

@@

Donors show increased willingness to donate to NGOs facing government restrictions/criticism while being financial transparent and accountable. While concerns about NGO mismanagement and lack of financial transparency may signal concerns about an NGO's operations and its ability to abide by the host country's regulatory environment, the presence of such transparency and accountability may instead convince donors that government targeting of NGOs may be ill-intentioned rather than simply a manifestation of financial oversight. Donors may view anti-NGO laws and inhospitable civic environments as a sign of governments looking to crack down on groups that seek to keep governments accountable, rather than merely auditing and punishing groups that violate routine regulations.

Therefore, both states and individual donors expand and contract their regulatory environments (for states) and support (for donors) for NGOs in response to the interplay between domestic politics and organizational characteristics. The same NGO feature that can increase the likelihood of government restriction can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

@@

::: {.content-visible when-profile="nice"}
\clearpage
:::

@@

At the time of designing the study, Charity Navigator---a large database of US nonprofit financial information---categorized international nonprofit activities into four general causes: (1) development, (2) humanitarian relief, (3) international affairs and human rights advocacy, and (4) environment.[^cn] Accordingly, we used four nonprofits that are stereotypical for each cause: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment). We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

@@

@@

We do not include any respondent-level covariates beyond the treatment variables. Because this is an experimental design, any statistical confounding is accounted for during the process of randomization and covariates should have no systematic effect on treatment effects. We do not work with the raw results of the model directly (though a complete table of results is included in Table A4). Given the conjoint design, we instead create a complete balanced grid of all 576 combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding) and use the model to calculate predicted probabilities of choice selection for each row of possible treatment values. We then collapse this set of predicted probabilities into estimated marginal means (EMMs) for specific features of interested while marginalizing or averaging over all other predicted variables [@LeeperHoboltTilley:2020; @BremerBurgisser:2023]. We report the causal effect of each manipulated feature using the average marginal component effect (AMCE), which is equivalent to the difference in estimated marginal means for specific feature levels. For example, when estimating the marginal means of the transparency treatment, we find the average predicted probability across the 288 rows of the reference grid where transparency is true and across the 288 rows where transparency is false---the difference between these two estimated marginal means is the AMCE, or the causal effect of the treatment holding all other values constant, on the probability scale.

@@

We present the posterior distributions of the marginal means and AMCEs for each of our experimental conditions in @fig-all-results and provide posterior medians, credible intervals, and other model diagnostics in Table A4 and Table A7. Because AMCEs are relative statements (i.e. contrasts between one feature level and a reference level), we try to use logical reference levels: for binary treatments like transparency and accountability, we calculate the difference between false and true levels; for ordered treatments like host country relationship, we calculate the differences between different levels of crackdown compared to no crackdown. To avoid imposing an artificial order on other unordered treatment variables, we report both marginal means and AMCEs [@LeeperHoboltTilley:2020].

@@

To contextualize the magnitude of the causal effects for our hypotheses of interest, we begin our analysis with a brief overview of the effects of organizational characteristics that nonprofits have less direct control over: their brand name, their issue area, and their funding sources. Brand recognition appears to be a powerful heuristic for donor decision making. Respondents were substantially more likely to prefer an organization when it was identified as the Red Cross, with a median posterior marginal mean of `r mms$red_cross$pct` (95% cred. int. = `r mms$red_cross$nice_ci`)---ten percentage points higher than the 25% probability that would be expected if respondents selected an organization at random. When using Amnesty International as the reference category, the AMCE for the Red Cross is `r amce$red_cross$pp` percentage points (`r amce$red_cross$nice_ci_pd_prefix`). Other organizations see much smaller marginal means and AMCEs. Compared to Amnesty International, Greenpeace causes a small `r amce$greenpeace$pp` percentage point increase (`r amce$greenpeace$nice_ci_pd`) and Oxfam causes a small `r amce$oxfam$pp_abs` percentage point decrease (`r amce$oxfam$nice_ci_pd`) in the probability of selecting an organization. The Red Cross brand name heuristic is the strongest of all the experimental treatments and features, signifying the organization's brand power and goodwill among potential donors.

@@

There is also some evidence that these two effects are equivalent and perhaps interchangeable. @fig-transp-acc (and Table A5) shows the posterior marginal means for all four combinations of possible levels of transparency and accountability. When respondents see both as "no" simultaneously, the estimated marginal mean is lower than any of the individual treatments in @fig-all-results, at `r mms_ta$transparency_no_accountability_no$pct` (`r mms_ta$transparency_no_accountability_no$nice_ci`), or 9 percentage points lower than the baseline probability of 25%. When either transparency or accountability is set to "yes", the estimated marginal mean is essentially identical at `r mms$transparency_yes$pct` and `r mms$accountability_yes$pct`, respectively. When both treatments are set to "yes", the estimated marginal mean is `r mms_ta$transparency_yes_accountability_yes$pct` (`r mms_ta$transparency_yes_accountability_yes$nice_ci`), which is roughly the same as the overall estimated marginal mean for the Red Cross. Holding all other treatments constant, the effect of transparency and accountability practices on their own are generally the same, and when combined, the overall favorability of the organization increases substantially. This could indicate that donors do not care about which specific type of organizational practice an NGO engages in and that instead they are looking for some sort of signal that the organization is following best practices in transparency or accountability.

@@

In addition to these overall trends in preferences, we can measure the causal effect of moving from friendly NGO--government relations to a more antagonistic relationship. Using friendly relationship as the reference category, facing criticism by the host government causes a `r amce$criticized_by_government$pp_abs` percentage point reduction in the probability of selection (`r amce$criticized_by_government$nice_ci_pd`). Participants respond more strongly as the relationship becomes more conflictual and restricted---an NGO facing restrictions on its work sees a `r amce$under_government_crackdown$pp_abs` percentage point reduction (`r amce$under_government_crackdown$nice_ci_pd`) in support. We thus find strong evidence against H~2~: if NGOs face restrictions abroad, individual donors are *less* likely to donate to them. For context, the causal effect of facing restrictions is the same magnitude as the Red Cross branding effect, but in the opposite direction---donors seem to penalize NGOs facing criticism to the same extent that they reward the Red Cross. This result is surprising, but clarifies findings in previous research. In a similar vignette-based experiment, @ChaudhryHeiss:2021 find no substantial effect of legal restrictions on the probability of donation, and any substantive crackdown-related effects are moderated by other treatments like NGO issue area or funding source. This negative effect might also be related to the notion of donor efficacy. When deciding how to maximize the impact or marginal benefit of their individual donation, donors look for signals that their money will make a difference. Transparency and accountability practices, in addition to knowing that other individuals regularly support the organization, all act as heuristics of efficacy. An NGO facing criticism or restrictions abroad, on the other hand, may signal that potential donations could be used for legal fees, or signal that the NGO had perhaps done something to deserve the legal limitations it faces. These restrictions thus serve as a negative heuristic, signaling that donations might not be used as effectively as donors might hope.

@@

Financial transparency and accountability practices both have a positive (and likely interchangeable) effect on NGO favorability, while government criticism and restriction have a negative effect and discourage potential donors from donating. As seen in @fig-transp-crackdown (and Table A6), these heuristics also interact with each other.[^transp-only] The estimated posterior marginal means for different relationships with host governments move in the same direction regardless of whether an organization engages in transparency practices: organizations with no conflict are most preferred, while organizations facing restrictions are least preferred. Transparency practices in the presence of restrictions seem to have two general effects. First, engaging in transparency offsets most of the negative effect of facing hostile civic environments. The estimated marginal mean for an organization with friendly host government relationships and no transparency is `r mms_tc$no_friendly_relationship_with_government$pct`, which is equivalent to the estimated marginal mean for an organization facing government restrictions that does engage in transparency. On average, donors are indifferent to both situations---again, a marginal mean of 25% represents the probability of selecting an organization at random---but the presence of transparency shifts NGOs facing restrictions from negatively preferred to indifferent, while the absence of transparency shifts NGOs with friendly relationships from positively preferred to indifferent. From a practical perspective, this suggests that NGOs operating in hostile civic environments can focus on improving specific organizational practices to offset the negative signals that accompany their legal difficulties.

@@

Second, worsening host government relationships weaken the positive effect of transparency. On its own, as seen in @fig-all-results, engaging in transparency causes a `r amce$transparency_yes$pp_abs` percentage point increase in the probability of a respondent selecting an organization. The right panel of @fig-transp-crackdown (and Table A6) shows how the causal effect of transparency changes across different types of government relationships. These AMCEs represent the difference in the estimated posterior marginal means of the two levels of transparency in the left panel. The effect shrinks as relationships become more negative: under friendly conditions, the median posterior AMCE of transparency is `r amce_tc$friendly_relationship_with_government$pp` percentage points (`r amce_tc$friendly_relationship_with_government$nice_ci_pd`); when an NGO is criticized, the transparency effect is `r amce_tc$criticized_by_government$pp` percentage points (`r amce_tc$criticized_by_government$nice_ci_pd`); when an NGO faces hostile restrictions, the effect drops to `r amce_tc$under_government_crackdown$pp` percentage points (`r amce_tc$under_government_crackdown$nice_ci_pd`). The causal effect remains substantially positive regardless of the relationship---even under the worst conditions, engaging in transparency causes a 9 percentage point boost---suggesting that NGOs facing crackdown can still increase their favorability with donors by signalling their commitment to transparency and accountability.

@@

# Conclusion

Over recent years, governments globally have systematically restricted NGOs using legal means. This trend of closing civic space has important implications for local and international NGOs as well as donors. Recent research has analyzed the impact of this crackdown on public or official aid donors [@DupuyPrakash:2018; @Herrold:2020; @RightSpringmanWibbels:2022]. However, we know relatively little about how state repression of NGOs affects the preferences of foreign private donors [@ChaudhryHeiss:2021]. Unlike state donors, individual donors have different motivations for engaging in philanthropy, and they may not necessarily withdraw or reduce support for NGOs facing harassment abroad. This is important because private philanthropy towards organizations working in international affairs continues to grow despite inflation---from 2021 to 2022, individual giving to nonprofits in international affairs grew by over 10.9% in the US [@GivingUSA:2023].

@@

These results have two important practical implications for NGO operations, fundraising, and survivability at a time when many INGOs are dealing with hostile host governments. First, while  NGOs have little to no control over host government crackdowns over their organizations, they do have control over organizational practices transparency and accountability, and more importantly, public perception of these organizational characteristics. As a result, they may need to rely entirely on improving individual donor perceptions of organizational transparency and accountability and emphasizing the need for private donor funds at a time of shrinking civic space. This article thus provides insight into the importance of different framing or informational heuristics that can motivate such individual donor giving. In addition to voluntarily emphasizing or sharing this information with likely donors, nonprofits could also potentially brainstorm how to get more positive media exposure regarding these characteristics in restrictive civic environments.

>>>>

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2025-01-20"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
correspondence-prefix: "Correspondence concerning this article should be addressed to"
abstract: |
  Governments across the world have increasingly used laws to restrict the work of nonprofits, which has led to a reduction in public or official foreign aid directed towards these groups. Many international nonprofits, in response, have turned to individual donors to offset the loss of traditional fundings. What are individual donors' preferences regarding donating to legally besieged nonprofits abroad? We conducted a conjoint experiment on a nationally representative sample of likely donors in the US and found that learning about host government criticism and restrictions on nonprofits decreases individuals' preferences to donate to them. However, organizational features such as financial transparency and accountability can protect against this dampening effect. Our results have important implications both for understanding private international philanthropy and how nonprofits can better frame their fundraising appeals at a time when they are facing restrictive civic spaces and hostile governments abroad.
additional-info: |
  **Data Availability:** All data and replication code are publicly available at <https://github.com/andrewheiss/silent-skywalk>
keywords:
  - philanthropy
  - conjoint experiments
  - donor heuristics
  - repression
  - NGOs
  - civil society
  - nonprofits
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
word-count: "9,300"
---

@@

In the remainder of the paper, we provide an overview of current research on private philanthropic giving to international NGOs. We use these existing theories to hypothesize how organizational practices like financial transparency and accountability---in isolation and when interacted with other organizational features---influence individual donor behavior in an environment of closing civic space. We then detail our conjoint study methodology and discuss our findings. Finally, we use our results to offer insights for NGOs facing legal restriction abroad.

@@

Though individual donors would like to have some assurance that their resources will be used appropriately and organizations are not spending too much on overhead [@HungHagerTian:2023], individuals interested in supporting charities cannot thoroughly vet every aspect of the organization [@Tremblay-BoirePrakash:2017; @ChaudhryHeiss:2021]. To more easily judge financial transparency and accountability, donors may use ratings by watchdog organizations like Charity Navigator and GuideStar as heuristics to guide their decisions.[^cn-multiple] Research shows that initiatives to address information gaps by these organizations can substantially increase donations [@Sloan:2009; @GordonKnockNeely:2009; @Tremblay-BoirePrakash:2017]. We expect that donors will respond to signals that organizations engage in transparency and accountability practices:

@@

However, previous literature has established that individual private donors do not make the same considerations as official donors when deciding to donate [@DesaiKharas:2018; @ChaudhryHeiss:2021]. Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. When a potential recipient NGO faces legal crackdown in its host country, private individual donors can respond by (1) increasing their donations as a sign of support and solidarity, (2) decreasing their donations, punishing the NGO for doing something to run afoul of its host government, or (3) not considering the host country legal environment and making no change in their donation behavior. @ChaudhryHeiss:2021 find conflicting donor responses to foreign restrictions---while some prospective donors punish nonprofits for doing something "wrong" to incur host government restrictions, most tended to increase their support, stating that restricted NGOs are "doing good work in countries where it is tough for groups like them to operate and they need all the help they can get" [@ChaudhryHeiss:2021, 496]. @ChaudhryDotsonHeiss:2021 explore how different combinations of donor characteristics determine which response individuals are likely to take when NGOs face crackdown. They find that donors with longer experiences with the nonprofit sector and high levels of social trust---i.e. those who frequently volunteer, regularly donate to charity, and trust political institutions---are more likely to maintain their support for international NGOs that face criticism or crackdown abroad. Thus, anti-NGO laws may act as a heuristic to individuals donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and seek to restrict their work.

@@

> **H~2~**: If NGOs face legal restrictions abroad, then individual private donors will have a higher likelihood of supporting or donating to them.

@@

Donors show increased willingness to donate to NGOs facing government restrictions/criticism while being financial transparent and accountable. While concerns about NGO mismanagement and lack of financial transparency may signal concerns about an NGO's operations and its ability to abide by the host country's regulatory environment, the presence of such transparency and accountability may instead convince donors that government targeting of NGOs may be ill-intentioned rather than simply a manifestation of financial oversight. Donors may view anti-NGO laws and inhospitable civic environments as a sign of governments looking to restrict groups that seek to keep governments accountable, rather than merely auditing and punishing groups that violate routine regulations.

Therefore, both states and individual donors expand and contract their regulatory environments (for states) and support (for donors) for NGOs in response to the interplay between domestic politics and organizational characteristics. The same NGO feature that can increase the likelihood of government restriction can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness. Based on the importance of transparency and accountability for donors in general, we expect that legal restriction will interact with these practices and shape donor preferences:

> **H~3~**: If NGOs face legal restrictions abroad *and* are financially transparent, then individual private donors will have a higher likelihood of supporting or donating to them.

@@

@@

At the time of designing the study, Charity Navigator---a large database of US nonprofit financial information---categorized international nonprofit activities into four general causes: (1) development, (2) humanitarian relief, (3) international affairs and human rights advocacy, and (4) environment.[^cn] Accordingly, we used four nonprofits that are stereotypical for each cause: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment).[^real-names] We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

@@

[^real-names]: We used real organization names rather than hypothetical, fictionalized names to allow the experimental task to better mirror reality [@McDonald:2020] and to reduce respondent cognitive burden when reflecting on organization characteristics [@MajnemerMeibauer:2022].

@@

We do not include any respondent-level covariates beyond the treatment variables. Because this is an experimental design, any statistical confounding is accounted for during the process of randomization and covariates should have no systematic effect on treatment effects. We do not work with the raw results of the multinomial model directly. Given the conjoint design, we instead create a complete balanced grid of all 576 combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding) and use the model to calculate predicted probabilities of choice selection for each combination of possible treatment values. We then collapse this set of predicted probabilities into estimated marginal means (EMMs) for specific features of interested while marginalizing or averaging over all other predicted variables [@LeeperHoboltTilley:2020; @BremerBurgisser:2023]. This marginalization process allows us to isolate the statistical effect of each feature in isolation. We include a complete table of model results in Table A4, along with a brief illustration of converting from regression coefficients to estimated marginal means.

We report the causal effect of each manipulated feature using the average marginal component effect (AMCE), which is equivalent to the difference in estimated marginal means for specific feature levels. For example, when estimating the marginal means of the transparency treatment, we find the average predicted probability across the 288 rows of the reference grid where transparency is true and across the 288 rows where transparency is false, holding all other experimental features and individual covariates constant---the difference between these two estimated marginal means is the AMCE, or the causal effect of the treatment on the probability scale.

@@

We present the posterior distributions of the marginal means and AMCEs for each of our experimental conditions in @fig-all-results and provide posterior medians, credible intervals, and other model diagnostics in Tables A6 and A8. Because AMCEs are relative statements (i.e. contrasts between one feature level and a reference level), we try to use logical reference levels: for binary treatments like transparency and accountability, we calculate the difference between false and true levels; for ordered treatments like host country relationship, we calculate the differences between different levels of crackdown compared to no crackdown. To avoid imposing an artificial order on other unordered treatment variables, we report both marginal means and AMCEs [@LeeperHoboltTilley:2020].

@@

To contextualize the magnitude of the causal effects for our hypotheses of interest, we begin our analysis with a brief overview of the effects of more foundational organizational characteristics: their organization name, their issue area, and their funding sources. Brand recognition appears to be a powerful heuristic for donor decision making. Respondents were substantially more likely to prefer an organization when it was identified as the Red Cross, with a median posterior marginal mean of `r mms$red_cross$pct` (95% cred. int. = `r mms$red_cross$nice_ci`)---ten percentage points higher than the 25% probability that would be expected if respondents selected an organization at random. When using Amnesty International as the reference category, the AMCE for the Red Cross is `r amce$red_cross$pp` percentage points (`r amce$red_cross$nice_ci_pd_prefix`). Other organizations see much smaller marginal means and AMCEs. Compared to Amnesty International, Greenpeace causes a small `r amce$greenpeace$pp` percentage point increase (`r amce$greenpeace$nice_ci_pd`) and Oxfam causes a small `r amce$oxfam$pp_abs` percentage point decrease (`r amce$oxfam$nice_ci_pd`) in the probability of selecting an organization. The Red Cross brand name heuristic is the strongest of all the experimental treatments and features, signifying the organization's brand power and goodwill among potential donors.

@@

There is also some evidence that these two effects are equivalent and perhaps interchangeable. @fig-transp-acc (and Table A7) shows the posterior marginal means for all four combinations of possible levels of transparency and accountability. When respondents see both as "no" simultaneously, the estimated marginal mean is lower than any of the individual treatments in @fig-all-results, at `r mms_ta$transparency_no_accountability_no$pct` (`r mms_ta$transparency_no_accountability_no$nice_ci`), or 9 percentage points lower than the baseline probability of 25%. When either transparency or accountability is set to "yes", the estimated marginal mean is essentially identical at `r mms$transparency_yes$pct` and `r mms$accountability_yes$pct`, respectively. When both treatments are set to "yes", the estimated marginal mean is `r mms_ta$transparency_yes_accountability_yes$pct` (`r mms_ta$transparency_yes_accountability_yes$nice_ci`), which is roughly the same as the overall estimated marginal mean for the Red Cross. Holding all other treatments constant, the effect of transparency and accountability practices on their own are generally the same, and when combined, the overall favorability of the organization increases substantially. This could indicate that donors do not care about which specific type of organizational practice an NGO engages in and that instead they are looking for some sort of signal that the organization is following best practices in transparency or accountability.

@@

In addition to these overall trends in preferences, we can measure the causal effect of moving from friendly NGO--government relations to a more antagonistic relationship. Using friendly relationship as the reference category, facing criticism by the host government causes a `r amce$criticized_by_government$pp_abs` percentage point reduction in the probability of selection (`r amce$criticized_by_government$nice_ci_pd`). Participants respond more strongly as the relationship becomes more conflictual and restricted---an NGO facing restrictions on its work sees a `r amce$under_government_crackdown$pp_abs` percentage point reduction (`r amce$under_government_crackdown$nice_ci_pd`) in support. We thus find strong evidence against H~2~: if NGOs face restrictions abroad, individual donors are *less* likely to donate to them. For context, the causal effect of facing restrictions is the same magnitude as the Red Cross branding effect, but in the opposite direction---donors seem to penalize NGOs facing criticism to the same extent that they reward the Red Cross. This result is surprising, but clarifies findings in previous research. In a similar vignette-based experiment, @ChaudhryHeiss:2021 find no substantial effect of legal restrictions on the probability of donation, and any substantive crackdown-related effects interact with and are dampened by other treatments like NGO issue area or funding source. This negative effect might also be related to the notion of donor efficacy. When deciding how to maximize the impact or marginal benefit of their individual donation, donors look for signals that their money will make a difference. Transparency and accountability practices, in addition to knowing that other individuals regularly support the organization, all act as heuristics of efficacy. An NGO facing criticism or restrictions abroad, on the other hand, may signal that potential donations could be used for legal fees, or signal that the NGO had perhaps done something to deserve the legal limitations it faces. These restrictions thus serve as a negative heuristic, signaling that donations might not be used as effectively as donors might hope.

@@

Financial transparency and accountability practices both have a positive (and likely interchangeable) effect on NGO favorability, while government criticism and restriction have a negative effect and discourage potential donors from donating. For our third hypothesis, we posit that these heuristics also interact with each other. This is confirmed in @fig-transp-crackdown (and Table A8)---the estimated posterior marginal means for different relationships with host governments move in the same direction regardless of whether an organization engages in transparency practices:[^transp-only] organizations with no conflict are most preferred, while organizations facing restrictions are least preferred. Transparency practices in the presence of restrictions seem to have two general effects. First, engaging in transparency offsets most of the negative effect of facing hostile civic environments. The estimated marginal mean for an organization with friendly host government relationships and no transparency is `r mms_tc$no_friendly_relationship_with_government$pct`, which is equivalent to the estimated marginal mean for an organization facing government restrictions that does engage in transparency. On average, donors are indifferent to both situations---again, a marginal mean of 25% represents the probability of selecting an organization at random---but the presence of transparency shifts NGOs facing restrictions from negatively preferred to indifferent, while the absence of transparency shifts NGOs with friendly relationships from positively preferred to indifferent. From a practical perspective, this suggests that NGOs operating in hostile civic environments can focus on improving specific organizational practices to offset the negative signals that accompany their legal difficulties.

@@

Second, worsening host government relationships weaken the positive effect of transparency. On its own, as seen in @fig-all-results, engaging in transparency causes a `r amce$transparency_yes$pp_abs` percentage point increase in the probability of a respondent selecting an organization. The right panel of @fig-transp-crackdown (and Table A8) shows how the causal effect of transparency changes across different types of government relationships. These AMCEs represent the difference in the estimated posterior marginal means of the two levels of transparency in the left panel. The effect shrinks as relationships become more negative: under friendly conditions, the median posterior AMCE of transparency is `r amce_tc$friendly_relationship_with_government$pp` percentage points (`r amce_tc$friendly_relationship_with_government$nice_ci_pd`); when an NGO is criticized, the transparency effect is `r amce_tc$criticized_by_government$pp` percentage points (`r amce_tc$criticized_by_government$nice_ci_pd`); when an NGO faces hostile restrictions, the effect drops to `r amce_tc$under_government_crackdown$pp` percentage points (`r amce_tc$under_government_crackdown$nice_ci_pd`). The causal effect remains substantially positive regardless of the relationship---even under the worst conditions, engaging in transparency causes a 9 percentage point boost---suggesting that NGOs facing crackdown can still increase their favorability with donors by signalling their commitment to transparency and accountability.

@@

# Discussion and conclusion

Over recent years, governments globally have systematically restricted NGOs using legal means. This trend of closing civic space has important implications for local and international NGOs as well as donors. Recent research has analyzed the impact of this crackdown on public or official aid donors [@DupuyPrakash:2018; @Herrold:2020; @RightSpringmanWibbels:2022]. However, we know relatively little about how state repression of NGOs affects the preferences of foreign private donors [@ChaudhryHeiss:2021]. Unlike state donors, individual donors have different motivations for engaging in philanthropy, and they may not necessarily withdraw or reduce support for NGOs facing harassment abroad. This is important because private philanthropy towards organizations working in international affairs continues to grow---following a decades-long trend, from 2021 to 2022, individual giving to nonprofits in international affairs grew by nearly 11% in the US [@GivingUSA:2023].

@@

These results have two important practical implications for NGO operations, fundraising, and survivability at a time when many INGOs are dealing with hostile host governments. First, while NGOs have little to no control over host government crackdowns over their organizations, they do have control over organizational practices transparency and accountability, and more importantly, public perception of these organizational characteristics. As a result, they may need to rely entirely on improving individual donor perceptions of organizational transparency and accountability and emphasizing the need for private donor funds at a time of shrinking civic space. This article thus provides insight into the importance of different framing or informational heuristics that can motivate such individual donor giving. In addition to voluntarily emphasizing or sharing this information with likely donors, nonprofits could also potentially brainstorm how to get more positive media exposure regarding these characteristics in restrictive civic environments. Moreover, we find that nonprofit transparency and accountability practices share the same causal effect, implying that donors might see these as interchangable concepts.
